Paper figures

### 2021-04-05 notebooks/17-PRS_dendrogram_cluster_investigation-020921.Rmd
unstashed some CSHL plotting edits

```{r}
library(reticulate)
#use_python("/home/oma21/miniconda3/envs/enm/bin/python")
use_condaenv('enm')
setwd('../')
#source('src/R/plot_density_prs.R')
source('src/R/get.subclusters.dendrogram.R')
library(RColorBrewer)
col = colorRampPalette(brewer.pal(n=9,name='Set1'))(10)
sensor_colors = col[1:8]
effector_colors = col[9:10]
```

@@

# Import goatools functions from src/utils.py
utils <- import('src.utils')
create_goa = utils$create_goea(gaf = 'data/raw/ontology/sgd.gaf', obo_fname='data/raw/ontology/go-basic.obo', background='data/interim/costanzo_gc_bg.tsv', sgd_info_tab = 'data/raw/ontology/SGD_features.tab')
goea = create_goa[[1]]
geneid2name = create_goa[[2]]
#utils$query_goatools(sensors_pcc,goea,geneid2name)
eff_go_terms <- list()
sens_go_terms <- list()
for(i in 1:4){
    print(i)
    subcluster <- effector_pcc%>%filter(clusters==i)
    query <- utils$query_goatools(subcluster,goea,geneid2name)
    eff_go_terms[i] = query[1,4]
}
effector_pcc <- unlist(eff_go_terms)%>%as.data.frame()%>%rownames_to_column()%>%mutate(rowname=as.numeric(rowname))%>%rename(cluster_go_term='.')%>%right_join(effector_pcc,by=c('rowname'='clusters'))
for(i in 1:4){
    print(i)
    subcluster <- sensors_pcc%>%filter(clusters==i)
    query <- utils$query_goatools(subcluster,goea,geneid2name)
    sens_go_terms[i] = query[1,4]
}
sensors_pcc <- unlist(sens_go_terms)%>%as.data.frame()%>%rownames_to_column()%>%mutate(rowname=as.numeric(rowname))%>%rename(cluster_go_term='.')%>%right_join(sensors_pcc,by=c('rowname'='clusters'))

>>>>

```{r}
library(reticulate)
#use_python("/home/oma21/miniconda3/envs/enm/bin/python")
use_condaenv('enm')
setwd('../')
#source('src/R/plot_density_prs.R')
source('src/R/get.subclusters.dendrogram.R')
library(RColorBrewer)
col =brewer.pal(n=8,name='Set1')
sensor_colors = col[1:4]
effector_colors = col[5:8]
```

@@

# Import goatools functions from src/utils.py
utils <- import('src.utils')
create_goa = utils$create_goea(gaf = 'data/raw/ontology/sgd.gaf', obo_fname='data/raw/ontology/go-basic.obo', background='data/interim/costanzo_gc_bg.tsv', sgd_info_tab = 'data/raw/ontology/SGD_features.tab')
goea = create_goa[[1]]
geneid2name = create_goa[[2]]
#utils$query_goatools(sensors_pcc,goea,geneid2name)
eff_go_terms <- list()
sens_go_terms <- list()
for(i in 1:4){
    print(i)
    subcluster <- effector_pcc%>%filter(clusters==i)
    query <- utils$query_goatools(subcluster,goea,geneid2name)
    print(head(query))
    eff_go_terms[i] = query[1,4]
}
effector_pcc <- as.character(eff_go_terms)%>%as.data.frame()%>%rownames_to_column()%>%mutate(rowname=as.numeric(rowname))%>%rename(cluster_go_term='.')%>%right_join(effector_pcc,by=c('rowname'='clusters'))%>%rename(clusters=rowname)

for(i in 1:4){
    print(i)
    subcluster <- sensors_pcc%>%filter(clusters==i)
    query <- utils$query_goatools(subcluster,goea,geneid2name)
    print(query)
    sens_go_terms[i] = query[1,4]
}
sensors_pcc <- as.character(sens_go_terms)%>%as.data.frame()%>%rownames_to_column()%>%mutate(rowname=as.numeric(rowname))%>%rename(cluster_go_term='.')%>%right_join(sensors_pcc,by=c('rowname'='clusters'))%>%rename(clusters=rowname)

### 2021-04-05 notebooks/paper_plots_0216_pcc_only_clustered.Rmd
unstashed some CSHL plotting edits

```{r}
library(reticulate)
# use_python("/home/oma21/miniconda3/envs/enm/bin/python")
use_condaenv("enm")
setwd("../")
source("src/R/plot_density_prs.R")
```

@@

```{r}
library(RColorBrewer)
col <- colorRampPalette(brewer.pal(n = 9, name = "Set1"))(12)
sensor_colors <- col[1:8] # brewer.pal(n = 6, name = "Set1")
effector_colors <- col[9:12] # brewer.pal(n=3, name= "Set1_r")
```

```{python}
sensor_colors = r.sensor_colors
effector_colors = r.effector_colors
sensors_pcc = e_pcc.sensors_df
sensors_pcc['go_group'] = sensors_pcc['cluster_go_term']
effector_pcc = e_pcc.effectors_df
effector_pcc['go_group'] = effector_pcc['cluster_go_term']
effector_order = effector_pcc.groupby('go_group').eff.median().sort_values().index.tolist()
sensor_order = sensors_pcc.groupby('go_group').sens.median().sort_values().index.tolist()
```

@@

>>>>

```{r}
library(reticulate)
# use_python("/home/oma21/miniconda3/envs/enm/bin/python")
use_condaenv("enm")
setwd("../")
source("src/R/plot_density_prs.R")
source("src/R/get.subclusters.dendrogram.R")
```

@@

```{r}
library(RColorBrewer)
col <- brewer.pal(n = 8, name = "Set1")
sensor_colors <- col[1:4] # brewer.pal(n = 6, name = "Set1")
effector_colors <- col[5:8] # brewer.pal(n=3, name= "Set1_r")
```

```{python}
sensor_colors = r.sensor_colors
effector_colors = r.effector_colors
sensors_pcc = e_pcc.sensors_df
sensors_pcc['go_group'] = sensors_pcc['cluster_go_term']
sensors_pcc['go_group'] = ['None'  if pd.isna(i) else i for i in sensors_pcc['go_group'].tolist()]
effector_pcc = e_pcc.effectors_df
effector_pcc['go_group'] = effector_pcc['cluster_go_term']
effector_order = effector_pcc.groupby('go_group').clusters.median().sort_values().index.tolist()
sensor_order = sensors_pcc.groupby('go_group').clusters.median().sort_values().index.tolist()
#sensor_order.append(None)
```

@@

```{python}
import copy
import scipy.cluster.hierarchy as sch
prs_mat = e_pcc.prs_mat
quantile_threshold =  0.95
method = 'ward'
metric = 'seuclidean'
q99 = np.quantile(prs_mat, quantile_threshold)
prs_mat_cl = copy.deepcopy(prs_mat)
prs_mat_cl[prs_mat_cl > q99] = q99
prs_mat_cl_orig = prs_mat_cl
# , optimal_ordering=True)
#row_linkage_cl = sch.linkage(sch.distance.pdist(prs_mat_cl), method=method)
# ,optimal_ordering=True)
#col_linkage_cl = sch.linkage(sch.distance.pdist(prs_mat_cl.T), method=method)

row_dist_orig = sch.distance.pdist((prs_mat),metric = metric)
col_dist_orig = sch.distance.pdist((prs_mat.T),metric = metric)
row_dist_orig_sqr = sch.distance.squareform(row_dist_orig)
col_dist_orig_sqr = sch.distance.squareform(col_dist_orig)

row_dist = sch.distance.pdist(prs_mat_cl,metric = metric)
col_dist = sch.distance.pdist(prs_mat_cl.T,metric = metric)

row_dist_sqr = sch.distance.squareform(row_dist)
col_dist_sqr = sch.distance.squareform(col_dist)

```{r}
cluster_method = 'ward.D2'
dist_row = as.dist(py$row_dist_sqr)
hclust_row = hclust(dist_row, cluster_method)
dist_col = as.dist(py$col_dist_sqr)
hclust_col = hclust(dist_col, cluster_method)
dend_row  <- as.dendrogram(hclust_row)
dend_col  <- as.dendrogram(hclust_col)
#effector_pcc <- get.subclusters.dendrogram(dend_row, effector_pcc)
#sensors_pcc <- get.subclusters.dendrogram(dend_col, sensors_pcc%>%select(-rowname))
library(dendextend)
#png('reports/figures/dendrogram021721/row_dendrogram_colored.png',width=800,height=800)
dend_row_plot <- dend_row[[2]] %>% 
    set("branches_k_color", k = 4,value = effector_colors[c(4,2,3,1)]) %>% 
    set("labels", NA) %>%    # change labelsk
    as.ggdend()%>%
    ggplot()#    plot(xlab="", sub="")}
#dev.off()
#png('reports/figures/dendrogram021721/col_dendrogram_colored.png',width=800,height=800)
dend_col_plot <-     dend_col[[1]] %>% 
    set("branches_k_color", k = 4,value = sensor_colors[c(3,2,4,1)]) %>% 
    set("labels", NA) %>%    # change labelsk
    as.ggdend()%>%
    ggplot()#    plot(xlab="", sub="")}
#dev.off()
```

```{python}
from matplotlib.lines import Line2D
pos = e_pcc_old.graph_gc.nodes('pos')
fig, ax = plt.subplots(figsize=(12,12))
legend_elements = [    ]
e_pcc.plot_network_spring(ax=ax,
                          node_size=1,
                          node_color='black',
 #                        node_size = [100 if i in sensors_pcc.orf_name.values or i in effector_pcc.orf_name.values else 1 for i in e_pcc.nodes],
                         #node_color = ['red' if i in sensors_pcc.orf_name.values else 'blue' if i in effector_pcc.orf_name.values else 'black' for i in e_pcc.nodes],
                         edge_color='black')
#                         node_shape=['^' if i in sensors_pcc.orf_name.values else 'v' if i in effector_pcc.orf_name.values else 'o' for i in e_pcc.nodes])
nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=sensors_pcc.orf_name.values, node_size=200, pos=pos,
                          node_color='black',
                          node_shape='^',edgecolors='black',
                          linewidths=1)
for itr, i in enumerate(sensor_order):
    orf_names_to_plot = sensors_pcc.loc[sensors_pcc.go_group==i, 'orf_name'].tolist()
    nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=orf_names_to_plot, node_size=200, pos=pos,
                          node_color=sensor_colors[itr],
                          node_shape='^',edgecolors='black',
                          linewidths=1)
    legend_elements.append(
        Line2D([0], [0], marker='^', color='black', label=f'Sensors ({i})',
                              markerfacecolor=sensor_colors[itr], markersize=12, linestyle="None")
    )
nx.draw_networkx_edges(nx.induced_subgraph(e_pcc.graph_gc, sensors_pcc.orf_name.tolist()), ax=ax , pos=pos, edge_color='red', alpha=0.5)
ax.set_facecolor('white')
legend_elements.extend(
    [Line2D([0], [0], marker='o', color='black', label='Genes',
                              markerfacecolor='black', markersize=4, linestyle="None"),
#                    Line2D([0], [0], marker='o', color='black', label='Effectors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
#                    Line2D([0], [0], marker='^', color='black', label='Sensors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
                       Line2D([0], [0], marker='o', color='black', label= 'High functional similarity',
                              markerfacecolor='black', markersize=0, linestyle="-"),
                   Line2D([0], [0], marker='o', color='red', label= 'Sensor-Sensor edges',
                              markerfacecolor='#018571', markersize=0, linestyle="-")]
)
lgd = ax.legend(handles=legend_elements, fontsize=22,loc='center left', bbox_to_anchor=(1.1, 0.5))
plt.savefig(f'{r.figure_folder}tmp_nw_sens.png',bbox_inches='tight')

```{python}
from matplotlib.lines import Line2D
pos = e_pcc_old.graph_gc.nodes('pos')
fig, ax = plt.subplots(figsize=(12,12))
legend_elements = [    ]
e_pcc.plot_network_spring(ax=ax,
                          node_size=1,
                          node_color='black',
 #                        node_size = [100 if i in sensors_pcc.orf_name.values or i in effector_pcc.orf_name.values else 1 for i in e_pcc.nodes],
                         #node_color = ['red' if i in sensors_pcc.orf_name.values else 'blue' if i in effector_pcc.orf_name.values else 'black' for i in e_pcc.nodes],
                         edge_color='black')
#                         node_shape=['^' if i in sensors_pcc.orf_name.values else 'v' if i in effector_pcc.orf_name.values else 'o' for i in e_pcc.nodes])
nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=effector_pcc.orf_name.values, node_size=100, pos=pos,
                      node_color='black',
                      node_shape='o')
for itr, i in enumerate(effector_order):
    orf_names_to_plot = effector_pcc.loc[effector_pcc.go_group==i,'orf_name'].tolist()
    nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=orf_names_to_plot, node_size=200, pos=pos,
                          node_color=effector_colors[itr],
                          node_shape='o',edgecolors='black',
                          linewidths=1)
    legend_elements.append(
            Line2D([0], [0], marker='o', color='black', label=f'Effectors ({i})',
                                  markerfacecolor=effector_colors[itr], markersize=12, linestyle="None")
        )
ax.set_facecolor('white')
legend_elements.extend(
    [Line2D([0], [0], marker='o', color='black', label='Genes',
                              markerfacecolor='black', markersize=4, linestyle="None"),
#                    Line2D([0], [0], marker='o', color='black', label='Effectors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
#                    Line2D([0], [0], marker='^', color='black', label='Sensors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
                       Line2D([0], [0], marker='o', color='black', label= 'High functional similarity',
                              markerfacecolor='black', markersize=0, linestyle="-"),
                   Line2D([0], [0], marker='o', color='blue', label= 'Effector-Effector edges',
                              markerfacecolor='#a6611a', markersize=0, linestyle="-")]
)
lgd = ax.legend(handles=legend_elements, fontsize=22,loc='center left', bbox_to_anchor=(1.1, 0.5))

nx.draw_networkx_edges(nx.induced_subgraph(e_pcc.graph_gc, effector_pcc.orf_name.tolist()), ax=ax , pos=pos, edge_color='blue',alpha=0.5)
plt.savefig(f'{r.figure_folder}tmp_nw_eff.png',bbox_inches='tight')

from src.visualize.visualize import heatmap_annotated
#import src
#import importlib
importlib.reload(src.visualize.visualize)
heatmap_annotated(prs_mat, prs_mat_cl_orig, figure_path=r.figure_folder, row_linkage=e_pcc.row_linkage, col_linkage=e_pcc.col_linkage, save_figure=True)#,
#                 row_colors=colors, col_colors=colors)

```{r}
library(coin)
df <- py$e_pcc$df%>%mutate(eff_group = as.factor(ifelse(orf_name%in%effector_pcc$orf_name,'eff','other')),
                    sens_group = factor(ifelse(orf_name%in%sensors_pcc$orf_name,'sens','other'),levels=c('sens','other')),
                    eff_norm = eff / nrow(.),
                    sens_norm = sens/ nrow(.))
pval_eff = pvalue(independence_test(eff~eff_group,df))
pval_sens = pvalue(independence_test(sens~sens_group,df))

eff_group_box <- df%>%ggplot(aes(x=eff_group,y=eff_norm))+
    geom_boxplot()+
  geom_signif(comparisons = list(c("eff", "other")), 
              map_signif_level=TRUE, y_position = 0.006)+
    xlab('')+
    ylab('Effectiveness')+
    scale_x_discrete(labels=c('Effectors','Others'))+
    theme_bw()
sens_group_box <- df%>%ggplot(aes(x=sens_group,y=sens_norm))+
    geom_boxplot()+
  geom_signif(comparisons = list(c("sens", "other")), 
              map_signif_level=TRUE, y_position = 0.0075)+
    xlab('')+
    ylab('Sensitivity')+
    scale_x_discrete(labels=c('Sensors','Others'))+
    theme_bw()+ylim(0,0.008)
#ggsave('tmp.png',sens_group_box)
heatmap_img <- glue('{figure_folder}prs_heatmap.png')%>%
    magick::image_read()
heatmap_img <- cowplot::ggdraw()+
    cowplot::draw_image(heatmap_img, width = 1)
eff_nw_img <- glue('{figure_folder}tmp_nw_eff.png') %>%
  magick::image_read() # %>%
eff_nw_img <- cowplot::ggdraw() +
  cowplot::draw_image(eff_nw_img, width = 1)
sens_nw_img <- glue('{figure_folder}tmp_nw_sens.png') %>%
  magick::image_read() # %>%
sens_nw_img <- cowplot::ggdraw() +
  cowplot::draw_image(sens_nw_img, width = 1)
cowplot::plot_grid(plot_grid(heatmap_img, plot_grid(sens_group_box, eff_group_box+ylim(0,0.007), nrow=2,labels=c('B','C')), nrow=1, rel_widths = c(0.7,0.3),labels=c('A','')),
                   plot_grid(dend_col_plot, sens_nw_img, nrow=1, rel_widths=c(0.2,.8), labels=c('D','E')),
                   plot_grid(dend_row_plot, eff_nw_img, nrow=1, rel_widths=c(0.2,.8), labels=c('F','G')), nrow=3, rel_heights=c(.5,.25,.25))
ggsave(glue('{figure_folder}fig1.pdf'),height=10,width=8)
